A cautionary note on the use of Ornstein Uhlenbeck and other models in macroevolutionary studies

### 2015-04-21 Manuscript/ohyou.tex

% Could move this down to the section on recommendations
    $\alpha$ can more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:

    \noindent
    This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
    If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.

  % I've left the actual details of the lit review in supp mat because I'm not usre how necessary they are here.
  \section{Uses of OU models}
  The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material). % Need to add ref to supp mat
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}.

    \begin{figure}[h]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year. See Supplementary Material for details %linkto supp mat
      }
      \label{figure.literature}
    \end{figure}

@@

  OU models are also commonly used to control for phylogeny in correlative analyses (Tables \ref{table.uses}; Supplementary Material).  
  Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
  This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data. 
  The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
  This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.

@@

    \begin{enumerate}
      \item $\alpha$ has a hard bound at zero, but no upper bound, meaning that the Maximum Likelihood profile is asymmetric. As values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small. This suggests that estimates of $\alpha$ are likely to be biased
      % make sure we really mean biased here
      , and that the asymptotic assumptions of Likelihood ratio tests on $\alpha$ are likely to be violated.
      \item blah
      \item measurement errors?
    \end{enumerate}

@@

    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was used to generate the data. 
    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains $>0.05$ even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
    This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure ; see Supplemental Material). % link to supp mat

    \begin{figure}[h]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/NTaxa.png}
      \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014. Two studies with \morethan 3000 taxa have been omitted for clarity. See Supplementary Material for details %linkto supp mat
      }
      \label{figure.ntaxa}
    \end{figure}

% Could put in something about interpreting parameters here?

\section{Effects of measurement error on OU model performance}
  Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep{} % add citation to Matt's model adequacy stuff
  Therefore, we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.

@@

    Table \ref{tables:yule_error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error. The expectation is that the OU model should fit better because the branch length transformation partially captures the non-Brownian component (the error). There are two points worth noting. First, the frequency with which the OU model is favoured increases with tree size. With as little as 5\% error, the OU model becomes extremely difficult to reject, even for trees with just 100 species. This is important for the interpretation of the OU model. We cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for. Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently $>$ 1. Large values of $\alpha$ are similarly difficult to interpret because they are indicative that the signal of the past has been overwritten.

\section{Prospects/ recommendations}
  \subsection{Brief discussion on methods and limitations for including measurement error}

  Recommendations

Interpreting results
 - OU is favoured over BM. If alpha is small it's basically BM. What
is alpha? Talk about half lives. add in ives and garland advice on what is big and what is small here?
 - Not stabilising selection
 - Not an issue if you're just mopping up error structure for PGLS (?)
 - If really want to be certain OU over BM you need 200 species
 - Account for measurement error

\section{Bayesian methods}
 (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)

\subsection{Bayesian simulations} % keep? or move it all to supp mat?
    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
    The choice of priors for OU models has not been fully explored. 
    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20. 
    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs. S2-S5). % add links to this
    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
    Multiple independent chains were run for each analysis to ensure convergence was reached. 
    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded) 
    We treated Bayes factors $>$ 2 as evidence favouring the OU model.

\section{Just an OU problem?}
Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}

@@

\section{Concluding comments/Conclusions}
\section{Epilogue - vaguely reproducible science}

%--------------------------------------
% I've left all this as is for now as we can slot relevant paragraphs into the appropriate places

\section{Discussion}

% This came from the intro but fits better later
%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes. $\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood. It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$. Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously. \citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989). They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes). This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.

Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways. Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes. But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation. The results we have presented here illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used.

In the case of the OU model, we can make a series of specific recommendations for future analyses:
\begin{enumerate}[(i)]
  \item At least 200 species should be included. The results of the simulations indicate that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases.  
  \item Likelihood ratio tests are untrustworthy. We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. The simulations indicate that likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered. 
  \item Simulate fitted models and compare these to your empirical results. A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values. These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}. This is important because we have shown that the shape of a phylogeny has consequences for parameter biases and hypothesis tests. Any given tree will therefore generate unique parameter estimates. Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your}. Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias. 
  \item Consider plausible alternative hypotheses. The results indicate that a simple parsimonious model can be mistaken for an OU process: namely a BM process in which a small amount of error is added to the data. Small ($<$ 1) values of $\alpha$ are quite compatible with this interpretation of the data. The effects of error become more severe with increasing tree size. 
\end{enumerate}

In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}. It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}. However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling. This is because the variances will themselves be estimates and subject to error. Treating such variances as error-free can potentially create biases (Freckleton unpub). If the variances are unknown then Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} can be used to estimate the non-Brownian component of the trait variance. However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.

Our review of the literature revealed that the OU model is frequently described and interpreted as a model of ‘stabilizing selection’. The OU model is attractive because it sets effective bounds in probability on the size of species’ traits, but to use the term ‘stabilizing selection’ is inaccurate and misleading. As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche. Under this formulation, α can be considered as the strength of the pull towards the primary optima \citep{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.

In terms of interpretation, it is worth pointing out that data will exhibit few signs of the limits to trait values that result from an OU process. This is because according to the OU process traits move back towards the optimum as their values become particularly large or small. Past history is effectively wiped over, as pointed out above, with the consequence that covariances between species become small (Fig. \ref{figure.traitsim}). Indeed, Figure \ref{figure.traitsim} implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present. This interpretation is particular important for inference of phylogenetic niche conservatism. Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}. A more parsimonious explanation would be that there is measurement error in the data. This seems particularly likely for climatic niche traits that are often averaged across an entire species range. Concluding from data on extant species that traits are limited by an OU-like process will therefore always be a leap of faith.

If ancillary data are available, however, this may not be the case. If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}.  A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances in equation (6) depend on both the shared distances between species and the distance of a node to the nearest tip. This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. These implementations should not be used with fossil data.

The results we report are not unique to the OU model. As noted above, \citet{grafen1989phylogenetic} pointed out that his $\rho$ parameter would likely be biased, and other transformations \citep[e.g., $\lambda$, $\delta$, ACDC;][]{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa} will exhibit similar behavior \citep[e.g., see][]{freckleton2002phylogenetic}. \citet{freckleton2002phylogenetic} explored the behavior of Pagel’s $\lambda$ and found that this statistic was biased but in manner that was likely to be conservative in the rejection of the Brownian model (also see Supplemental Fig. S7). Some transformations may be unbiased: for example, Pagel’s $\kappa$ is a transformation of branch lengths, and is essentially the “standard diagnostic” that is used in testing the assumptions of phylogenetic contrasts, and known to behave appropriately \citep[Supplemental Fig. S6;][]{garland1992procedures}.

In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species. The OU model is a good example of this: large values of $\alpha$ will lead to a loss of history and the dataset will only represent the most recent evolutionary changes. Moreover, different processes can easily yield similar patterns \citep[e.g.,][]{Revell:2008ab} with the consequence that rejection of the Brownian model in favour of another does not necessarily say anything about process. This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}. We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation. The results presented above highlight that in the case of the OU model, this is especially important.

  \subsection{Other evolutionary models}
    Although our focus is on the OU model we also explored three other commonly used models of trait evolution -- $\kappa$, $\lambda$, and $\delta$ \citep{Pagel:1997aa,Pagel:1999aa} -- fitted using Maximum Likelihood with the same simulated data. 
    These models are designed to test for speciational evolution ($\kappa$), strength of phylogenetic signal ($\lambda$) and accelerating or decelerating evolution ($\delta$). 
    The point of these additional tests is to assess the extent to which parameter behaviors vary across models and the results are reported in the Supplementary Material.

@@

\bibliographystyle{biolinnsoc}
\bibliography{ohyou}

\section{Appendix}
  \label{section:models}
  \setcounter{equation}{0}

@@

\input{ohyou_suppmat}

>>>>

  % I've left the actual details of the lit review in supp mat because I'm not usre how necessary they are here.
  \section{Uses of OU models}
  The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material). % Need to add ref to supp mat
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}. 
  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilizing selection'', this is inaccurate and misleading. As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche. 
  Under this formulation, $\alpha$ can be considered as the strength of the pull towards the primary optima \citep{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.

    \begin{figure}[h]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year. See Supplementary Material for details. %linkto supp mat
      }
      \label{figure.literature}
    \end{figure}

@@

  OU models are also commonly used to control for phylogeny in correlative analyses (Table \ref{table.uses}; Supplementary Material).  
  Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
  This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data. 
  The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
  This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. 
  Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.

@@

    \begin{enumerate}
      \item $\alpha$ has a hard bound at zero but no upper bound, meaning that the Maximum Likelihood profile is asymmetric. As values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small. This suggests that estimates of $\alpha$ are likely to be biased
      % make sure we really mean biased here
      , and that the asymptotic assumptions of Likelihood ratio tests on $\alpha$ are likely to be violated.
      \item blah
      \item measurement errors?
    \end{enumerate}

@@

    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was used to generate the data. 
    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains \textgreater 0.05 even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
    This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure \ref{figure:ntaxa}; see Supplemental Material). % link to supp mat

    \begin{figure}[h]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/NTaxa.png}
      \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014. Two studies with \textgreater 3000 taxa have been omitted for clarity. See Supplementary Material for details. %linkto supp mat
      }
      \label{figure:ntaxa}
    \end{figure}

    We provide recommendations related to these findings below. % name of section?

\section{Effects of measurement error on OU model performance}
  Measurement error can strongly effect the results of comparative analyses \cite{silvestro2015}, and appears to influence whether OU is the favoured model across a range of datasets \citep{}. % add citation to Matt's model adequacy stuff
  Therefore, we also investigated whether adding error to our simulated data influenced estimates of $\alpha$.

@@

    Table \ref{tables:yule_error} shows the proportion of data sets in which the OU model is favoured over the Brownian model for data simulated under Brownian motion with error. 
    The expectation is that the OU model should fit better because the branch length transformation partially captures the non-Brownian component (the error). There are two points worth noting. 
    First, the frequency with which the OU model is favoured increases with tree size. With as little as 5\% error, the OU model becomes extremely difficult to reject, even for trees with just 100 species. 
    This is important for the interpretation of the OU model. 
    We cannot conclude anything about evolutionary process from a single optimum OU model unless error is adequately accounted for. 
    Second, for moderate amounts of error (5-10\%), the ML estimates of $\alpha$ are consistently \textgreater 1. 
    Large values of $\alpha$ are similarly difficult to interpret because they are indicative that the signal of the past has been overwritten.

\section{Prospects: limitations, recommendations and potential solutions?}
  \subsection{Limitations of the OU model}
  Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways. 
  Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes. 
  But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation. 
  The results of our simulations illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used (Figure \ref{figure:ntaxa}, Supplementary Table S1).

  In the case of the OU model, there are several limitations worth highlighting:
\begin{enumerate}[(i)]
  \item \textbf{Type I error rates are high when sample size is low.} 
  The results of the simulations indicate that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases.  
  \item \textbf{Likelihood ratio tests are untrustworthy.}
  We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered. 
  \item \textbf{Measurement error increases Type I error rates.} 
  Our results indicate that a simple Brownian process can be mistaken for an OU process when a small amount of error is added to the data. 
  The effects of measurement error become more severe with increasing tree size.
\end{enumerate}

  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution. 
  This is particularly important for inference of phylogenetic niche conservatism. 
  Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}. 
  A more parsimonious explanation would be that there is measurement error in the data.
  This seems particularly likely for climatic niche traits that are often averaged across an entire species range. 
  Concluding from data on extant species that traits are limited by an OU-like process will therefore always be a leap of faith.

% Not sure this is needed anywhere
%In terms of interpretation, it is worth pointing out that data will exhibit few signs of the limits to trait values that result from an OU process. This is because according to the OU process traits move back towards the optimum as their values become particularly large or small. Past history is effectively wiped over, as pointed out above, with the consequence that covariances between species become small (Fig. \ref{figure.traitsim}). Indeed, Figure \ref{figure.traitsim} implies that the outcome of the OU process is indistinguishable from an acceleration of the rate of evolution towards the present.

\subsection{Potential solutions and recommendations}
  Although there are limitations with the OU model, there are also several ways researchers can get around these problems as follows:

  \begin{enumerate}[(i)]
    \item \textbf{Use at least 200 species.} 
    If you really want to be confident that the OU model is favoured over the Brownian model, you need to fit the models to more than 200 species. 
    Note that this does not solve the issue of measurement error. in the data.
    \item \textbf{Simulate fitted models and compare these to your empirical results.}
    A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values. These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}. This is important because we have shown that the shape of a phylogeny has consequences for parameter biases and hypothesis tests. Any given tree will therefore generate unique parameter estimates. Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your}. Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias. 
    \item \textbf{Consider plausible alternative hypotheses.}
    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$parameter when an OU model is favoured. 
    We explain this in detail below.
    \item \textbf{Account for measurement error.} % details
    % This is from the original Discussion
    %In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}. It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}. However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling. This is because the variances will themselves be estimates and subject to error. Treating such variances as error-free can potentially create biases (Freckleton unpub). If the variances are unknown then Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} can be used to estimate the non-Brownian component of the trait variance. However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances. 
    \item \textbf{If possible, include ancillary data.}
    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}
    This is implemented in GEIGER \citep{Harmon:2008aa}.
    A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances (see Appendix \ref{equation:OUcov} depend on both the shared distances between species and the distance of a node to the nearest tip. 
    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). 
    Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. 
    These implementations should not be used with fossil data.
    \item \textbf{Use a Bayesian approach.} See below. %Or cite our preprint?
\end{enumerate}

  Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
  Thus this does not reflect any kind of OU process underlying the data.

  To avoid misinterpreting results in these cases, we recommend inspecting the $\alpha$ parameter values as well as looking at which model is favoured. 
  If $\alpha$ is small, then this should not be interpreted as any OU-like process (including phylogenetic niche conservatism etc.).

  $\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:

  \noindent
  This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
  If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
  We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
  However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process.

% add in ives and garland advice on what is big and what is small here? Or earlier?

\subsection{Bayesian simulations} 
% (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)
    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
    The choice of priors for OU models has not been fully explored. 
    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20. 
    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs. S2-S5). % add links to this
    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
    Multiple independent chains were run for each analysis to ensure convergence was reached. 
    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded) 
    We treated Bayes factors $>$ 2 as evidence favouring the OU model.

\section{Just an OU problem?}
%Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}

%The modification of Brownian models to account for non-Brownian processes can be traced back to \citet{grafen1989phylogenetic} who introduced a parameter $\rho$ to account for non-linear scaling of evolutionary changes. $\rho$ is a power transformation of node heights in the phylogeny, and is fitted by maximum likelihood. It is worth noting that Grafen warned explicitly about potential drawbacks with $\rho$. Specifically: (i) the parameter is likely to be biased; (ii) its asymptotic sampling variance is non-zero; and (iii) values of the parameter should be interpreted cautiously. \citet{freckleton2002phylogenetic} showed that $\rho$ was indeed biased as predicted by Grafen (1989). They also showed that another parameter, Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa}, also exhibited biases (but that these were likely to have minimal consequences in analyses using reasonable sample sizes). This strongly suggests that simulations are required to test the properties of such parameters, and that as recommended by \citet{grafen1989phylogenetic} “careless” parameterization of the phylogenetic variance-covariance matrix should be avoided.

@@

 %The results we report are not unique to the OU model. As noted above, \citet{grafen1989phylogenetic} pointed out that his $\rho$ parameter would likely be biased, and other transformations \citep[e.g., $\lambda$, $\delta$, ACDC;][]{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa} will exhibit similar behavior \citep[e.g., see][]{freckleton2002phylogenetic}. \citet{freckleton2002phylogenetic} explored the behavior of Pagel’s $\lambda$ and found that this statistic was biased but in manner that was likely to be conservative in the rejection of the Brownian model (also see Supplemental Fig. S7). Some transformations may be unbiased: for example, Pagel’s $\kappa$ is a transformation of branch lengths, and is essentially the “standard diagnostic” that is used in testing the assumptions of phylogenetic contrasts, and known to behave appropriately \citep[Supplemental Fig. S6;][]{garland1992procedures}.

\section{Concluding comments/Conclusions}
% Conclusion from original
%In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species. The OU model is a good example of this: large values of $\alpha$ will lead to a loss of history and the dataset will only represent the most recent evolutionary changes. Moreover, different processes can easily yield similar patterns \citep[e.g.,][]{Revell:2008ab} with the consequence that rejection of the Brownian model in favour of another does not necessarily say anything about process. This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}. We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation. The results presented above highlight that in the case of the OU model, this is especially important.

\section{Epilogue: (vaguely) reproducible science}
At the symposium that generated this special issue, one of us (NC) gave a talk on reproducibility in science. 
We have therefore tried to make this paper as reproducible as possible. 
All simulated phylogenies, data, and R code are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.
However, although our R code is provided, it's messy and difficult to use. 
We also do not provide an automated way of reproducing the data collection for our literature review.
Nor do we use tools like Travis CI, Docker, or packrat etc.
% Add stuff about BayesTraits here. 
We included this epilogue to highlight that reproducibility is really hard, but we are trying to improve! 
With a little effort most people should be able to produce something vaguely reproducible, and move us all slightly closer to truly reproducible science.

@@

% Bibliography
\bibliographystyle{biolinnsoc}
\bibliography{ohyou}

% Appendix
\section{Appendix}
  \label{section:models}
  \setcounter{equation}{0}

@@

% Supplementary Material
\input{ohyou_suppmat}
